---
title: "Using the queuecomputer package"
author: "Anthony Ebert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the queuecomputer package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The purpose of the package queuecomputer is to compute, deterministically, the output of a queue network given the arrival and service times for all customers. The most important functions are `queue_step`, `lag_step` and `wait_step`. 

# Input format

The first argument to the functions `queue_step`, `lag_step` and `wait_step` is a vector of arrival times. For example:

```{r}
library(queuecomputer)
library(dplyr)

arrivals <- cumsum(rexp(100))

head(arrivals)

service <- rexp(100)

departures <- queue_step(arrivals = arrivals, service = service)

str(departures,1)
```

# Resourcing schedule

The resourcing schedule is specified with either a non-zero natural number, a `server.stepfun` or a `server.list` object. Use a non-zero natural number when the number of servers does not change over time. The `server.stepfun` specifies a step function to indicate how many servers are available throughout the day. The computation speed for `queue_step()` is much faster when using a `server.stepfun` rather than a `server.list` input for the `servers` argument. 

We create a `server.stepfun` object with the `as.server.stepfun` function. 

```{r}
# Zero servers available before time 10
# One server available between time 10 and time 50
# Three servers available between time 50 and time 100
# One server available from time 100 onwards
resource_schedule <- as.server.stepfun(c(10,50,100), c(0, 1, 3, 1))

resource_schedule

departures <- queue_step(arrivals = arrivals, service = service, servers = resource_schedule)

str(departures,1)
```

The `server.list` object is a list of step functions which represent each server, the range is $\{0,1\}$, where 0 represents unavailable and 1 represents available and the knots represent the times where availability changes. 

The `as.server.list` function is used to create a `server.list` object.

```{r}
# Server 1 is available before time 10.
# Server 2 is available between time 15 and time 30.
# Server 3 is available after time 10. 
as.server.list(list(10, c(15,30), 10), c(1,0,0))
```

# Setting up a queue network

It is simple to set up a chain of queueing elements with `queuecomputer`. Suppose passengers must walk to a queue, then wait for service and then wait for their bags. 

```{r}
library(queuecomputer)
library(dplyr)

set.seed(500)

n <- 100

arrivals <- cumsum(rexp(n))
service_l <- rexp(n, 0.8)
service_q <- rexp(n, 0.5)
arrivals_b <- cumsum(rexp(n, 0.8))

# The queue elements can be computed one by one. 

departures_1 <- lag_step(arrivals, service_l)
departures_2 <- queue(departures_1, service = service_q, servers = 2)
departures_3 <- wait_step(departures_2, arrivals_b)

# Or the queue elements can be chained together with the %>% operator. 

departures <- lag_step(arrivals, service_l) %>% queue_step(service = service_q, servers = 2) %>% wait_step(arrivals_b)

all(departures == departures_3)

# Plot densities for this tandem queueing network

colours <- rainbow(4)
plot(density(arrivals, from = 0), 
  col = colours[1], xlim = c(0, 220), ylim = c(0, 0.015), 
  main = "Density plot")
lines(density(departures_1, from = 0), col = colours[2])
lines(density(departures_2, from = 0), col = colours[3])
lines(density(departures_3, from = 0), col = colours[4])
legend(150,0.012, legend = c("Start walk",
    "Finish walk",
    "Finish service", 
    "Pick up bag"),
    col = colours, lwd = 1, cex = 0.8
)



```



